\Huge Conferences in the Era of Expensive Carbon

### 2019-06-24 main.tex
Polishing

Our first proposal is simple: {\bf Every ACM-sponsored conference should
publicly report its carbon footprint.} Most conferences'
footprints will be dominated by participants' air travel, but the accounting
should go beyond this to include air travel to in-person PC meetings and
estimated hotel and food CO2e footprints.

Public accounting will nudge conference organizers and attendees to change
their behavior. By analogy, chain restaurants in the US are required by law
to post calorie counts of food items on menu boards, and studies show that
enlightened customers order, on average, up to 50 fewer calories a day as a
result~\cite{menu}.

@@

SIGPLAN has also explored how to mitigate a conference's CO2e emissions by
purchasing \emph{carbon offsets}~\cite{CarbonOFfsetReport}.
%
A carbon offset is sold by a vendor who uses the raised funds to finance an
activity that permanently avoids emitting some amount of greenhouse gases.
The veracity and permanence of this CO2e emission avoidance is certified by
a watchdog organization. (Typically, planting trees is not a certifiable
activity, since it's hard to guarantee the trees won't be cut down, but
installing methane capture devices on landfill sites or buying
fuel-efficient stoves to replace open-fire cooking in poor communities can
be.)
%
Many organizations, including companies such as Google, Dell, Microsoft,
General Motors, Delta Airlines, Lyft, and Expedia, numerous universities,
and some academic societies, now buy carbon offsets in order to reduce their
net footprint. ACM conferences can do the same, and the purchases can be
included in the public accounting we are proposing.  In fact, ACM recently
decided that conferences could offer attendees the option to pay for an
offset during registration; the conference then uses the designated funds to
purchase the offset from a qualified vendor.  Some conferences are also
experimenting with corporate sponsorship of bulk offsets for the whole
conference.\footnote{We acknowledge that carbon offsets are a somewhat
  controversial 
  tool~\cite{anderson2012inconvenient,
  carbon-offsets-are-not-our-get-out-jail-free-card}.  Besides issues with
  verifiability and permanence, some observers worry that---especially given
  their currently artificially low cost---they will encourage complacency,
  leading organizations to postpone making the difficult changes that will
  be required to sustainably decarbonize society.  We believe that offsets
  should be considered as a useful short-term expedient, to be considered in
  parallel with, not instead of, more difficult cutting.} \mwh{Should
  we mention the FCRC numbers here?}

Beyond encouraging conferences to take these sorts of simple steps, public
accounting of conference emissions will generate data that future organizers
can use.  To this end, emissions data should be published both in aggregate
form (total and per-participant footprint), for quick reference, and in raw
form (a suitably anonymized record for each individual attending each
conference, including originating city), and it should be made available in
a central place and in a uniform format, to facilitate analysis.

ACM should provide tools for gathering and publicizing this data. As a
start, SIGPLAN recently worked with ACM's SIG Governing Board to develop an
air-travel-focused CO2e calculator for conferences. Users can upload
conference registration data, and the calculator will estimate the CO2e cost
of air travel for those sufficiently far away. This calculator could be
expanded to account for other forms of conference emissions.

@@

What should ACM do with the funds collected from this surcharge?  One can
imagine many good uses, including carbon offsetting, 
defraying the costs of virtualizing conferences (livestreaming, etc.),
and supporting ``green''
computing research. For an example of the last: what if ACM funded a major cross-cutting
research initiative specifically aimed at understanding how to best replace
or approximate the socializing and networking aspects of conferences in a
virtual setting? We would hope that the cost of carbon pricing and the funding of
initiatives such as this will stimulate creative rethinking of the conference model
itself---for example, seriously considering alternatives such as
rapid-turnaround journal-only publishing models, once-a-year
megaconferences, and entirely virtual conferences~\cite{NCN}.
%
A potential sticking point here is that some of these strategies will also
significantly reduce conference revenues; in the limit, this may impact the
income stream of ACM itself, which could make emissions reduction
politically unpalatable unless ACM's conference-focused business model is
also adjusted.  Planning for this eventuality should begin now!

>>>>

Our first proposal is simple: {\bf Every ACM-sponsored conference should
publicly report its carbon footprint.} 
%
Public accounting will nudge conference organizers and attendees to change
their behavior. (By analogy, chain restaurants in the US are required by law
to post calorie counts of food items on menu boards, and studies show that
enlightened customers order, on average, up to 50 fewer calories a day as a
result~\cite{menu}.)

@@

SIGPLAN has also explored how to mitigate a conference's CO2e emissions by
purchasing \emph{carbon offsets}~\cite{CarbonOFfsetReport}.
%
A carbon offset is sold by a vendor who uses the raised funds to finance an
activity that permanently avoids emitting some amount of greenhouse gases.
The veracity and permanence of this CO2e emission avoidance is certified by
a watchdog organization. (Typically, planting trees is not a certifiable
activity, since it's hard to guarantee the trees won't be cut down, but
installing methane capture devices on landfill sites or buying
fuel-efficient stoves to replace open-fire cooking in poor communities can
be.)
%
Many organizations, including companies such as Google, Dell, Microsoft,
General Motors, Delta Airlines, Lyft, and Expedia, numerous universities,
and some academic societies, now buy carbon offsets in order to reduce their
net footprint. ACM conferences can do the same, and the purchases can be
included in the public accounting we are proposing.  In fact, ACM recently
decided that conferences could offer attendees the option to pay for an
offset during registration; the conference then uses the designated funds to
purchase the offset from a qualified vendor.  Some conferences are also
experimenting with corporate sponsorship of bulk offsets for the whole
conference.\footnote{We acknowledge that carbon offsets are a 
  controversial 
  tool~\cite{anderson2012inconvenient,
  carbon-offsets-are-not-our-get-out-jail-free-card}.  Besides issues with
  verifiability and permanence, some observers worry that---especially at
  their current artificially low cost---they will encourage complacency,
  leading organizations to postpone making the difficult changes that will
  be required to sustainably decarbonize society.  We believe that offsets
  should be considered as a useful short-term expedient, to be considered in
  parallel with, not instead of, more difficult cutting.}
% \mwh{Should
% we mention the FCRC numbers here? 45 out of 2,700 used CO2e offsets}
% That's getting too much down into the details, I think -- also, these
% numbers reflect a hastily organized, preliminary version of what we really
% want to do.

Beyond encouraging conferences to take these sorts of simple steps, public
accounting of conference emissions will generate data that future organizers
can use.  To this end, emissions data should be published both in aggregate
form (total and per-participant footprint), for quick reference, and in raw
form (a suitably anonymized record for each individual attending each
conference, including originating city), and it should be made available in
a central place and in a uniform format, to facilitate analysis.  (Most
conferences footprints will be dominated by participants' air travel, but the
data gathered should go beyond this to include air travel to in-person PC
meetings and estimated hotel and food CO2e footprints.)

ACM should provide tools for gathering and publicizing this data. As a
first step, SIGPLAN recently worked with ACM's SIG Governing Board to develop an
air-travel-focused CO2e calculator for conferences. Users can upload
conference registration data, and the calculator will estimate the CO2e cost
of air travel for those sufficiently far away. This calculator could be
expanded to account for other forms of conference emissions.

@@

What should ACM do with the funds collected from this surcharge?  One can
imagine many good uses, including carbon offsetting, defraying the costs of
virtualizing conferences (livestreaming, etc.), and supporting ``green''
computing research. As an example of the last, ACM could help fund a major
cross-cutting research initiative specifically aimed at understanding how to
best replace or approximate the socializing and networking aspects of
conferences in a virtual setting?

Beyond simple belt-tightening, carbon pricing and the funding of initiatives
such as this should stimulate creative rethinking of the conference model
itself---for example, seriously considering alternatives such as
rapid-turnaround journal-only publishing models, once-a-year
megaconferences, and entirely virtual conferences~\cite{NCN}.
%
A potential sticking point, however, is that some of these strategies will also
significantly reduce conference revenues; in the limit, this may impact the
income stream of ACM itself, which could make emissions reduction
politically unpalatable unless ACM's conference-focused business model is
also adjusted.  Planning for this eventuality should begin now!
